---
title: "Bayesian Fundamentals Analysis Task"
author: "Scott Sun"
date: "1/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("https://github.com/reyesem/reyesem.github.io/raw/master/files/MA483/MA483Setup.R")
```

Load the data as follows.
```{r}
receipts <- tibble(
  Vehicle = 1:25,
  `Hours Billed` = c(2, 1, 3, 2, 1,
                     1, 4, 1, 2, 4,
                     3, 1, 2, 5, 2,
                     3, 2, 1, 3, 2,
                     5, 6, 2, 5, 1))
```

### Graphical Summary

```{r, fig.width=5, fig.asp=0.6}
ggplot(data = receipts,
       mapping = aes(x = `Hours Billed`)) +
  geom_bar(aes(y = ..prop..), fill = "orange") +
  labs(x = "Hours Billed",
       y = "Proportion",
       title = "Receipts' Hours Billed") +
  theme_bw()
```

### Parameter / Model
The length of time $Y \sim Exp(\lambda)$. Let the number of hours billed $C = \lceil Y \rceil$.

The parameter of interest is $\lambda$, the average number of hours a vehicles would remain in the garage.

We want to study $Pr(\frac{1}{\lambda} > 2|\boldsymbol{y})$.

### Prior Distribution

Given that
$$
\begin{aligned}
1 &= \int_{0}^{\infty} \pi(\lambda) d\lambda \\
  &= \int_{0}^{\infty} \frac{\Gamma(a+b)}{\Gamma(a)\Gamma(b)} e^{-a\lambda}(1-e^{-\lambda})^{b-1} d\lambda \\
  &= \int_{0}^{\infty} \frac{\Gamma(a+b)}{\Gamma(a)\Gamma(b)} (e^{-\lambda})^{a-1} (1-e^{-\lambda})^{b-1} \cdot e^{-\lambda} d\lambda
\end{aligned}
$$

Let $\theta = e^{-\lambda}$.
$$1 = \int_{0}^{1} \frac{\Gamma(a+b)}{\Gamma(a)\Gamma(b)} \theta^{a-1}(1-\theta)^{b-1}d\theta$$
Thus, $\theta \sim Beta(a,b)$

Now, solve the hyperparameters by the following system of equations.
$$
\begin{cases}
Pr(\frac{1}{\lambda} < 5) = Pr(\theta < e^{-1/5}) = 0.95 \\
\\
Pr(\frac{1}{\lambda} > 1/2) = Pr(\theta > e^{-2}) = 0.95 \\
\end{cases}
$$

```{r}
fn <- function(params) {
  a <- params[1]
  b <- params[2]
  
  eq1 <- pbeta(exp(-1/5), shape1 = a, shape2 = b)
  eq2 <- 1 - pbeta(exp(-2), shape1 = a, shape2 = b)
  
  (0.95-eq1)^2 + (0.95-eq2)^2
}

optim.result1 <- optim(c(1,1), fn)
optim.result1$par %>%
  round(3)
```

$$
\begin{aligned}
a &= 2.212 \\
b &= 2.506
\end{aligned}
$$

### Posterior

Before solve the posterior distribution, we shall derive the likelihood.

$$
\begin{aligned}
f_C(c|\lambda)
&=Pr(C = c|\lambda)  \\
&= Pr(\lceil Y \rceil = c|\lambda) \\
&= Pr(c-1 < Y \leq c|\lambda) \\
&= \int_{c-1}^{c} \lambda e^{-\lambda y} dy \\
&= (e^{\lambda} - 1)e^{-\lambda c} = (\frac{1}{\theta} - 1)\theta^c \\
&= f_C(c|\theta)
\end{aligned}
$$

Now,
$$
\begin{aligned}
\pi(\theta|\boldsymbol{c}) 
&\propto f(\boldsymbol{c}|\theta) \cdot \pi(\theta) \\
&\propto (\frac{1}{\theta} - 1)^n \theta^{n\bar{c}} \cdot \theta^{a-1}(1 - \theta)^{b-1} \\
&\propto \theta^{n\bar{c}+a-1} (1-\theta)^{b-1}(\frac{1}{\theta}-1)^n \cdot \frac{\theta^n}{\theta^n}\\
&\propto \theta^{n\bar{c}-n+a-1}(1-\theta)^{n+b-1}
\end{aligned}
$$
Therefore, $\theta|\boldsymbol{c} \sim Beta(n\bar{c}-n+a, n+b)$.

### Conclusion
Since the question of interest is $Pr(\frac{1}{\lambda} > 2|\boldsymbol{c})$,

$$
\begin{aligned}
Pr(\frac{1}{\lambda} > 2|\boldsymbol{c}) 
&= Pr(\theta > e^{-1/2}|\boldsymbol{c}) \\
&= 1 - \int_{0}^{e^{-1/2}}\pi(\theta|\boldsymbol{c})d\theta
\end{aligned}
$$

```{r}
a <- optim.result1$par[1]
b <- optim.result1$par[2]
n <- length(receipts$`Hours Billed`)
c_bar <- mean(receipts$`Hours Billed`)

post_alpha <- n*c_bar-n+a
post_beta <- n+b

(1 - pbeta(exp(-1/2), shape1 = post_alpha, shape2 = post_beta)) %>%
  round(3)
```

Given the prior knowledge and the data provided, the probability that vehicles on average remain parked for more than 2 hours during an event is 0.461.

