---
title: "Bayesian Fundamentals Homework"
author: "Scott Sun"
date: "1/13/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("Bayesian Initialization Script.R")
```

## Concepts
### Problem 1 (Predicting Enrollment)
No. The question of interest is to find $Pr(Y^* > 250|\boldsymbol{y})$. The probability can be obtained by
$$Pr(Y^* > 250|\boldsymbol{y}) = 1 - \sum_{y^* = 0}^{250}\pi(y^*|\boldsymbol{y})$$
where $\pi(y^*|\boldsymbol{y})$ is the density of the posterior predictive ditribution.

### Problem 2 (Predicting Enrollment, Cont)
Based on the prior that $\theta \sim Beta(a,b)$, $Pr(H_0) = 0$ since Beta distribution is continuous. Besides, the posterior $\theta|\boldsymbol{y}$ would be continuous as well. As the result, $Pr(H_0|\boldsymbol{y}) = 0$. In this scenario, the test is meaningless because we've already believed $H_0$ is impossible and no amount of data would change our mind.

### Problem 4 (Thesis Completion, Cont)
Since $\theta$ is not defined yet, support of $T$ is unknown. Therefore, we need to include indicator function $\mathbb{I}(t_{(1)}>\theta)$ so that the posterior density should be
$$
\pi(\lambda,\theta|\boldsymbol{t}) \propto
\pi(\lambda,\theta) \cdot \lambda^ne^{\lambda n \bar{t} + n \lambda \theta} \cdot
\mathbb{I}(t_{(1)}>\theta)
$$

## Computations
### Problem 6 (Lake Level)
Solve the following system of equations.
$$
\begin{cases}
Pr(576<\mu<580) = 0.95 \\
\\
Pr(578<\mu<582) = 0.95 \\
\end{cases}
$$

```{r, warning=FALSE}
fn1 <- function(params) {
  theta <- params[1]
  tau <- params[2]
  
  eq1 <- pnorm(580, mean = theta, sd = tau) - pnorm(576, mean = theta, sd = tau)
  eq2 <- pnorm(582, mean = theta, sd = tau) - pnorm(578, mean = theta, sd = tau)
  
  (eq1 - 0.95)^2 + (eq2 - 0.95)^2
}

optim.result1 <- optim(c(579, 1), fn1)
par <- optim.result1$par %>%
  round(3)
names(par) <- c("theta", "tau")
par
```

\pagebreak

### Problem 8 (Lake Level, Cont)
Load the data as follows.
```{r}
huron <- tibble( 
  Year = seq(1875, 1972), 
  level = c(580.38, 581.86, 580.97, 580.80, 579.79, 580.39, 580.42, 580.82, 
            581.40, 581.32, 581.44, 581.68, 581.17, 580.53, 580.01, 579.91, 
            579.14, 579.16, 579.55, 579.67, 578.44, 578.24, 579.10, 579.09, 
            579.35, 578.82, 579.32, 579.01, 579.00, 579.80, 579.83, 579.72, 
            579.89, 580.01, 579.37, 578.69, 578.19, 578.67, 579.55, 578.92, 
            578.09, 579.37, 580.13, 580.14, 579.51, 579.24, 578.66, 578.86, 
            578.05, 577.79, 576.75, 576.75, 577.82, 578.64, 580.58, 579.48, 
            577.38, 576.90, 576.94, 576.24, 576.84, 576.85, 576.90, 577.79, 
            578.18, 577.51, 577.23, 578.42, 579.61, 579.05, 579.26, 579.22, 
            579.38, 579.10, 577.95, 578.12, 579.75, 580.85, 580.41, 579.96, 
            579.61, 578.76, 578.18, 577.21, 577.13, 579.10, 578.25, 577.91, 
            576.89, 575.96, 576.80, 577.68, 578.38, 578.52, 579.74, 579.31, 
            579.89, 579.96)) 
```

```{r}
theta <- 580
tau <- 3

post_mean <- (theta/tau^2 + sum(huron$level)/(1.5^2))/(1/tau^2 + length(huron$level)/(1.5^2))
post_sd <- 1/sqrt(1/tau^2 + length(huron$level)/(1.5^2))

post_nullProb <- pnorm(579.5, post_mean, post_sd) - pnorm(578.5, post_mean, post_sd)
post_altProb <- 1 - post_nullProb
prior_nullProb <- pnorm(579.5, theta, tau) - pnorm(578.5, theta, tau)
prior_altProb <- 1 - prior_nullProb

logBF01 <- log((post_nullProb/post_altProb)*(prior_altProb/prior_nullProb))
logBF01 %>%
  round(3)
```

### Problem 9 (Lake Levels, Alternative Perspective)
```{r}
post_alpha <- 0.001 + length(huron$level)/2
post_beta <- 0.001 + 0.5*sum((huron$level-579)^2)
```

Let $r$ and $s$ be two endpoints of the HDI and solve the following system of equations.
$$
\begin{cases}
\pi(\tau=r|\boldsymbol{y}) = \pi(\tau=s|\boldsymbol{y}) \\
\\
\int_r^s\pi(\tau|\boldsymbol{y}) = 0.98
\end{cases}
$$

```{r}
fn2 <- function(endpts){
  r <- endpts[1]
  s <- endpts[2]
  
  eq1 <- dgamma(r, shape = post_alpha, rate = post_beta) -
    dgamma(s, shape = post_alpha, rate = post_beta)
  
  eq2 <- pgamma(s, shape = post_alpha, rate = post_beta) -
    pgamma(r, shape = post_alpha, rate = post_beta)
  
  (eq1)^2 + (0.98 - eq2)^2
}

optim.result2 <- optim(c(qgamma(0.025, shape = 1008.47, rate = 1180.2), 
                         qgamma(0.975, shape = 1008.47, rate = 1180.2)),
                       fn2)
optim.result2$par %>%
  round(3)
```
Therefore, the 98% HDI for $\tau$ is (0.399, 0.783).

## Complete Analysis
### Problem 10 (Oncology)
Given that $Y \sim Bino(20, \theta),~\theta \sim Beta(a, b)$

Now,
$$
\left\{
  \begin{array}{ll}
  E(\theta) = \frac{a}{a+b} = 0.9\\
  Pr(\theta < 0.8) = 0.05
  \end{array}
\right.
$$

The hyperparameters a and b in $\pi(\theta)$ can be solved as follows.
```{r}
fn_oncology <- function(params) {
  a <- params[1]
  b <- params[2]
  
  eq1 <- a/(a+b)
  eq2 <- pbeta(0.8, a, b)
  
  (eq1 - 0.9)^2 + (eq2 - 0.05)^2
}

optim.result_oncology <- optim(c(9, 1), fn_oncology)
optim.result_oncology$par %>%
  round(3)
```

The posterior density can be obtained by
$$
\begin{aligned}
\pi(\theta|y) 
&\propto \pi(\theta)\cdot f(y|\theta)\\
&= \frac{\Gamma(a+b)}{\Gamma(a)\Gamma(b)} \cdot \theta^{a-1}(1-\theta)^{b-1} \cdot 
\binom{20}{y} \theta^y(1-\theta)^{20-y}\\
&\propto \theta^{a+y-1}(1-\theta)^{20+b-y-1}
\end{aligned}
$$
which is the kernel of $Beta(a+y, 20+b-y)$.

The posterior mode can be solved as follows.
```{r}
n <- 20
y <- 14
a <- optim.result_oncology$par[1]
b <- optim.result_oncology$par[2]


fn_oncology_mode <- function(theta) {
  -dbeta(theta, shape1 = a+y, shape2 = n+b-y, log = TRUE)
}
optim.result_oncology_mode <- optim(0.2, fn_oncology_mode, method = "Brent", lower = 0, upper = 1)
optim.result_oncology_mode$par %>%
  round(3)
```

\pagebreak

### Problem 12 (Cereal Boxes)

```{r}
cereal <- tibble( 
  Amount = c(13.55, 12.51, 12.71, 12.51, 12.04, 12.40, 12.19, 
             13.74, 12.09, 13.35, 13.41, 12.95, 12.24, 12.64, 12.06)
  )
```

Since $y_i = \theta + \epsilon_i$, $y_i|\theta \sim Unif(\theta-1, \theta+1)$.

And $\theta \sim Unif(12,13)$

Now,
$$
\left\{
  \begin{array}{ll}
  f(\boldsymbol{y}|\theta) = (\frac{1}{2})^{15} \cdot \mathbb{I}(y_{(1)} > \theta-1, y_{(15)} < \theta+1)\\
  \pi(\theta) = \mathbb{I}(12 < \theta < 13)
  \end{array}
\right.
$$

```{r}
lower_bd <- max(c(12, max(cereal$Amount)-1))
upper_bd <- min(c(13, min(cereal$Amount)+1))
print(c(lower_bd, upper_bd))
```

The posterior density is given by
$$
\begin{aligned}
\pi(\theta|y) 
&\propto f(\boldsymbol{y}|\theta) \cdot \pi(\theta)\\ 
&= (\frac{1}{2})^{15} \cdot \mathbb{I}(y_{(1)} > \theta-1, y_{(15)} < \theta+1) \cdot
\mathbb{I}(12 < \theta < 13)\\
&\propto \mathbb{I}(y_{(1)} > \theta-1, y_{(15)} < \theta+1) \cdot \mathbb{I}(12 < \theta < 13)\\
&= \mathbb{I}(12.74 < \theta < 13)
\end{aligned}
$$
which is the kernel of $Unif(12.74, 13)$

```{r}
qunif(0.025, lower_bd, upper_bd)
qunif(0.975, lower_bd, upper_bd)
```

Therefore, the equal-tail 95% credible intervel is (12.75, 13.00).

