---
title: "Bayesian Computation Analysis Task"
author: "Scott Sun"
date: "2/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("https://github.com/reyesem/reyesem.github.io/raw/master/files/MA483/MA483Setup.R")
library(gridExtra)
library(gamlss.dist)
```

```{r}
fish <- read.csv("Fish.csv")
```

### Graphical Summary

```{r, fig.width=5, fig.asp=0.6}
ggplot(data = fish,
       mapping = aes(x = count)) +
  geom_bar(aes(y = ..prop..), fill = "orange") +
  labs(x = "The number of fish caught by a group",
       y = "Proportion",
       title = "Distribution of a the number of fish caught by a group") +
  theme_bw()
```

### Parameter and Model

Let $Y:=$ the number of fish caught by a group. Then $Y_i \stackrel{i.i.d.}{\sim}$ a Zero-Inflated Poisson Distribution, whose probability mass function is
$$
f(y | \lambda, \theta) = 
\begin{cases}
  \theta + (1-\theta)e^{-\lambda} & y = 0\\
  \\
  (1-\theta)\frac{\lambda^ye^{-\lambda}}{y!} & y = 1,2...
\end{cases}  
$$

Here, $\theta$ represents proportion of groups which choose not to fish, and $\lambda$ represents the average counts for those groups who fish.

Now, define $\theta_0 = 1 - \theta$ as the proportion of groups choosing to fish.

Since one of the objectives is to study the probability that a given fish group catch at least 2 fish, define
$$
\begin{aligned}
p 
&= \Pr(Y \geq 2) \\
&= 1 - (\Pr(Y=0) + \Pr(Y=1)) \\
&= 1 - (\theta + (1-\theta)e^{-\lambda}+(1-\theta)\lambda e^{-\lambda})
\end{aligned}
$$

The variables of interest in the problem are $\theta_0$, and p.

### Prior Distribution

Based on the domains of $\theta_0$ and $\lambda$, we assume the following prior distributions.
$$
\begin{aligned}
\theta_0 &\sim Beta(a, b) \\
\lambda &\sim Gamma(s, t) \\
\theta_0 &\perp\!\!\!\perp \beta
\end{aligned}
$$

Then, $\{a, b, s, t\}$ can be calculated as follows.

According to the spokesman, there roughly 75% $\pm$ 10% of groups choosing to fish.
```{r}
fn1 <- function(params) {
  a <- params[1]
  b <- params[2]
  
  eq1 <- a/(a+b)
  eq2 <- a*b/((a+b)^2*(a+b+1))
  
  (0.75 - eq1)^2 + (0.1 - eq2)^2
}

betaOptParam <- optim(c(5, 5), fn1)
betaOptParam <- betaOptParam$par %>%
  round(3)
names(betaOptParam) <- c("a", "b")
betaOptParam
```

Apart from that, groups who fish catch approximately 5 $\pm$ 1 fish.

```{r}
fn2 <- function(params) {
  s <- params[1]
  t <- params[2]
  
  eq1 <- s*t
  eq2 <- s*t^2
  
  (5 - eq1)^2 + (1 - eq2)^2
}

gammaOptParam <- optim(c(1, 1), fn2)
gammaOptParam <- gammaOptParam$par %>%
  round(3)
names(gammaOptParam) <- c("s", "t")
gammaOptParam
```


### Sample from the Posterior Distribution

```{stan output.var="StanFish", eval=FALSE}
data{
  int <lower = 0> N;
  int <lower = 0> y[N];
}

parameters{
  real <lower = 0, upper = 1> theta;
  real <lower = 0> lambda;
}

transformed parameters{
  real <lower = 0> p;
  real <lower = 0> theta_0;
  p = 1 - (theta + (1-theta)*exp(-lambda)*(1+lambda));
  theta_0 = 1 - theta;
}

model{
  target += beta_lpdf(theta_0 | 0.656, 0.219) 
              + gamma_lpdf(lambda | 25, 0.002);
  
  for (n in 1:N) {
    if (y[n] == 0)
      target += log_sum_exp(bernoulli_lpmf(1 | theta), 
                            bernoulli_lpmf(0 | theta) 
                              + poisson_lpmf(y[n] | lambda));
    else
      target += bernoulli_lpmf(0 | theta) 
                  + poisson_lpmf(y[n] | lambda);
  } 
}
```

```{r, eval=FALSE}
saveRDS(StanFish, "StanFish.rds")
```

```{r}
StanFish <- readRDS("StanFish.rds")
dataList.fish <- list(N = nrow(fish),
                      y = fish$count)
fit.fish <- stan(model_code = StanFish@model_code,
                 data = dataList.fish)
```

```{r, echo=FALSE}
fit.fish
```

```{r}
fishParams.df <- stan_to_df(fit.fish)
```

```{r, fig.asp=0.4}
p1 <- qplot(data = fishParams.df,
            x = fishParams.df$theta_0,
            geom = "density",
            fill = I("orange")) +
  labs(x = expression(theta[0] ~ "| y"),
       y = "Density",
       title = "Posterior Distribution")

p2 <- qplot(data = fishParams.df,
            x = fishParams.df$p,
            geom = "density",
            fill = I("cornflowerblue")) +
  labs(x = "p | y",
       y = "Density",
       title = "Posterior Distribution")

grid.arrange(p1, p2, ncol = 2)
```

### Posterior Predictive

```{r}
y_star <- rZIP(n = nrow(fishParams.df), 
               mu = fishParams.df$lambda,
               sigma = fishParams.df$theta)
y_star <- data.frame(y_star = y_star)
```

```{r, fig.width=5, fig.asp=0.6}
qplot(data = y_star,
      x = y_star,
      geom = "density",
      fill = I("brown2"),
      alpha = I(0.8)) + 
  labs(x = expression(y^"*"~"| y"),
       y = "Density",
       title = "Posterior Predictive Distribution")
```

### Conclusion

Therefore, according to the inference result above, the 95% credible interval of $\theta_0|\boldsymbol{y}$ is (0.37, 0.49); the 95% credible interval of $p|\boldsymbol{y}$ is (0.37, 0.49).

