---
title: "Comparing Groups AT"
author: "Scott Sun"
date: "2/17/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("https://github.com/reyesem/reyesem.github.io/raw/master/files/MA483/MA483Setup.R")
```

```{r}
mm <- tibble(
  Type  = c(rep("plain", 6), rep("peanut butter", 6)),
  Color = rep(c("red", "orange", "yellow", "green", "blue", "brown"), 2),
  Count = c(20, 12, 24, 26, 16, 9,
            11,  5,  9,  9,  7, 7)
)
```

### Graphical Summary 

```{r}
ggplot(data = mm,
       aes(x = Color, y = Count, fill = Type)) +
  geom_bar(stat = "identity", position="dodge") +
  scale_fill_manual(values = c("cadetblue1", "cadetblue")) +
  labs(x = "Color of the M&M bean",
       y = "Count",
       fill = "",
       title = "M&M beans color distribution in two different types of packet") +
  theme_bw()
```

Based on graphic above, the color distributions of two packets seem different. Besides, peanut butter packet has more beans than the plain packet

### Parameters and Model

Let $\boldsymbol{Y}_{1}$ be a vector of counts for the peanut butter packet.

Let $\boldsymbol{Y}_{2}$ be a vector of counts for the plain packet.

Let $\boldsymbol{\theta}_{j}$ be the probability distribution of color for j-type packet, where
$$\sum_{i=1}^{k}\boldsymbol\theta_{j}[i] = 1$$

$$
\begin{aligned}
\boldsymbol{Y}_1 &\sim Multinomial(\boldsymbol{\theta}_1) \\
\boldsymbol{Y}_2 &\sim Multinomial(\boldsymbol{\theta}_2) \\
\boldsymbol{Y}_1 &\perp\!\!\!\perp \boldsymbol{Y}_2
\end{aligned}
$$

The question of interest is to test the following hypothesis test:
$$
H_{0}:\boldsymbol{\theta}_1 = \boldsymbol{\theta}_2 ~~~~~~~ 
H_1: \boldsymbol{\theta}_1 \neq \boldsymbol{\theta}_2
$$

### Prior Distribution
Based on the property of $\boldsymbol{\theta}_j$, assume that
$$
\begin{aligned}
\boldsymbol{\theta}_j &\sim Dirichlet(\boldsymbol{\alpha}) \\
\boldsymbol{\theta}_1 &\perp\!\!\!\perp \boldsymbol{\theta}_2
\end{aligned}
$$

Apart from that, also assume $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 1$.

### Sample from Posterior distribution

Under $H_0$, let $\boldsymbol{\theta}_1 = \boldsymbol{\theta}_2 = \boldsymbol{\theta}$.
$$
\begin{aligned}
\pi(\boldsymbol{\Theta}|\boldsymbol{y}) 
&\propto f(\boldsymbol{y}|\boldsymbol{\Theta}) \cdot \pi(\boldsymbol{\Theta}) \\
&= f(\boldsymbol{y}_1|\boldsymbol{\theta})f(\boldsymbol{y}_2|\boldsymbol{\theta}) \cdot
\pi(\boldsymbol{\theta}) & \text{due to indep}
\end{aligned}
$$


Under $H_1$,
$$
\begin{aligned}
\pi(\boldsymbol{\Theta}|\boldsymbol{y}) 
&\propto f(\boldsymbol{y}|\boldsymbol{\Theta}) \cdot \pi(\boldsymbol{\Theta}) \\
&= f(\boldsymbol{y}_1|\boldsymbol{\theta}_1)f(\boldsymbol{y}_2|\boldsymbol{\theta}_2) \cdot
\pi(\boldsymbol{\theta}_1)\pi(\boldsymbol{\theta}_2) & \text{due to indep}
\end{aligned}
$$

```{stan output.var="StanMM0", eval = FALSE}
data {
  int<lower = 0> K;
  int<lower = 0> count_1[K];
  int<lower = 0> count_2[K];
  vector[K] alpha;
}

parameters{
  simplex[K] theta;
}

model{
  target += dirichlet_lpdf(theta | alpha);
  target += multinomial_lpmf(count_1 | theta) + 
          multinomial_lpmf(count_2 | theta);
}
```

```{stan output.var="StanMM1", eval = FALSE}
data {
  int<lower = 0> K;
  int<lower = 0> count_1[K];
  int<lower = 0> count_2[K];
  vector[K] alpha;
}

parameters{
  simplex[K] theta_1;
  simplex[K] theta_2;
}

model{
  target += dirichlet_lpdf(theta_1 | alpha) + 
          dirichlet_lpdf(theta_2 | alpha);
  target += multinomial_lpmf(count_1 | theta_1) + 
          multinomial_lpmf(count_2 | theta_2);
}
```

```{r, eval=FALSE}
saveRDS(StanMM0, "StanMM0.rds")
saveRDS(StanMM1, "StanMM1.rds")
```

```{r}
StanMM0 <- readRDS("StanMM0.rds")
StanMM1 <- readRDS("StanMM1.rds")

dataList.MM <- list(
  K = length(table(mm$Color)),
  count_1 = filter(mm, Type == "peanut butter") %>% 
    select(Count) %>% 
    unlist() %>% 
    unname(),
  count_2 = filter(mm, Type == "plain") %>% 
    select(Count) %>% 
    unlist() %>% 
    unname(),
  alpha = rep(1, 6)
    )

fit.mm0 <- stan(model_code = StanMM0@model_code,
                data = dataList.MM)
fit.mm1 <- stan(model_code = StanMM1@model_code,
                data = dataList.MM)
```

### Conclusion
```{r}
fit.mm0
```

```{r}
fit.mm1
```

```{r}
evidence.mm0 <- bridge_sampler(fit.mm0, silent = TRUE)
evidence.mm1 <- bridge_sampler(fit.mm1, silent = TRUE)
post_prob(evidence.mm0, evidence.mm1,
          model_names = c("H0", "H1")) %>%
  round(3)
```

Therefore, there is strong evidence that the color distributions of two packet are not different.
