---
title: "Regression AT"
author: "Scott Sun"
date: "2/18/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("Bayesian Initialization Script.R")
options(mc.cores = 3)
```

```{r}
avengers <- read.csv("Avengers.csv")
```

### Graphical Summary

```{r, fig.width=5, fig.asp=0.6}
avengers <- avengers %>%
  mutate(t = Weekend - 1)

qplot(data = avengers,
      x = t, y = Amount, color = Movie,
      geom = "point") +
  geom_line(aes(group = Movie)) +
  scale_color_manual(values = c("darkorchid4", "brown1")) +
  labs(x = "Number of weekends after the release",
       y = "Box office during weekend",
       title = "The decaying curve of box office during weekend") +
  theme_bw() +
  theme(legend.position = "bottom")
```

According to the graph above, the weekend earnings of both Avenger movies decayed exponentially.

### Parameter and Model

Let $A_t$ represent the amount the film earns during the i-th weekend after the release.

Assume the following model on $A_t$:
$$
\begin{aligned}
A_t &\sim N(\mu_t, \sigma^2) \\
\\
\mu_t &= \alpha_j\exp((\beta_0 + \beta_1\mathbb{I}(\text{End Game})_t)\cdot t)
\end{aligned}
$$
where
$$
\begin{aligned}
\alpha_{1} &:=  \text{amount earned by End Game in the first weekend}\\
\alpha_{2} &:=  \text{amount earned by Infinity War in the first weekend}\\
\beta_0,\beta_1 &< 0\end{aligned}
$$

### Prior Distribution

$$
\begin{aligned}
\sigma &\sim \text{half-Cauchy}(0, 1) \\
\beta_0, \beta_1 &\sim \text{half-Cauchy}(0, 1)
\end{aligned}
$$

### Sample from Posterior Distribution

```{stan output.var="StanAvengers0", eval=FALSE}
data{
  int<lower = 0> N;
  real<lower = 0> a[2];
  int<lower = 0, upper = 1> EndGame[N];
  real Amount[N];
  int Time[N];
}

parameters{
  real<upper = 0> beta0;
  real<lower = 0> sigma;
}

model{
  real mu;
  sigma ~ cauchy(0, 1);
  beta0 ~ cauchy(0, 1);
  
  for (i in 1:N) {
    if (EndGame[i] == 0)
      mu = a[1] * exp(beta0 * Time[i]);
    else
      mu = a[2] * exp(beta0 * Time[i]);
    
    Amount[i] ~ normal(mu, sigma);
  }
}
```


```{stan output.var="StanAvengers1", eval=FALSE}
data{
  int<lower = 0> N;
  real<lower = 0> a[2];
  int<lower = 0, upper = 1> EndGame[N];
  real Amount[N];
  int Time[N];
}

parameters{
  real<upper = 0> beta0;
  real<upper = 0> beta1;
  real<lower = 0> sigma;
}

model{
  real mu;
  sigma ~ cauchy(0, 1);
  beta0 ~ cauchy(0, 1);
  beta1 ~ cauchy(0, 1);
  
  for (i in 1:N) {
    if (EndGame[i] == 0)
      mu = a[1] * exp((beta0 + beta1 * EndGame[i])*Time[i]);
    else
      mu = a[2] * exp((beta0 + beta1 * EndGame[i])*Time[i]);
    
    Amount[i] ~ normal(mu, sigma);
  }
}
```

```{r, eval=FALSE}
saveRDS(StanAvengers0, "StanAvengers0.rds")
saveRDS(StanAvengers1, "StanAvengers1.rds")
```

```{r}
StanAvengers0 <- readRDS("StanAvengers0.rds")
StanAvengers1 <- readRDS("StanAvengers1.rds")
```

```{r}
a <- filter(avengers, t == 0) %>%
  select(Amount) %>%
  unlist() %>%
  unname()

avengers <- avengers %>%
  mutate(Movie = ifelse(Movie == "Endgame", 1, 0))

dataList.avengers <- list(N = nrow(avengers),
                          a = a,
                          EndGame = avengers$Movie,
                          Amount = avengers$Amount,
                          Time = avengers$t)
```

```{r}
fit.avengers0 <- stan(model_code = StanAvengers0@model_code,
                     data = dataList.avengers)
fit.avengers1 <- stan(model_code = StanAvengers1@model_code,
                     data = dataList.avengers)
```

```{r}
fit.avengers0
```

```{r}
fit.avengers1
```

\pagebreak

### Conclusion
```{r}
evidence.avengers0 <- bridge_sampler(fit.avengers0, silent = TRUE)
evidence.avengers1 <- bridge_sampler(fit.avengers1, silent = TRUE)
post_prob(evidence.avengers0, evidence.avengers1,
          model_names = c("H0", "H1")) %>%
  round(6)
```

Therefore, we have strong evidence that the decay rates are different. The weekend earnings of Avengers: Infinity War and Avengers: Endgame have different behavior.