---
title: "Statistical Process Analysis Task"
author: "Scott Sun"
date: "12/9/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("https://github.com/reyesem/reyesem.github.io/raw/master/files/MA483/MA483Setup.R")
```

Load the dataset and categorize eruptions lasting longer than 3 minutes as `long` and the rest of all as `short`. The head of the data frame is shown as the follows.

```{r}
data("faithful")
faithful <- faithful %>%
  mutate(grp = ifelse(eruptions > 3, "long", "short"))
head(faithful, 5) %>%
  knitr::kable()
```

## Graphics

```{r, fig.width=5, fig.asp=0.5}
ggplot(data = faithful,
       mapping = aes(x = eruptions, y = waiting)) +
  geom_point(aes(color = grp)) +
  labs(x = "Length of eruption / min",
       y = "Length of waiting / min",
       color = "",
       title = "Length of eruption vs. length of waiting") +
  theme_bw()
```

\pagebreak
Based on the graphic above, there exists two clusters in the observations, which align with the classification made at the beginning.

## Parameters & Process

Let $N \equiv \text{the total number of eruptions the visitor saw}$. $N \sim Geom(p)$, and $p$ is the probability of observing an eruption longer 3 minutes.

Then, let 
$$
\begin{aligned}
T &\equiv \text{total amount of time} \\
U_i &\equiv \text{length of the}~i^{th}~\text{eruption},~i=1,...,N-1 \\
W_i &\equiv \text{length of the}~i^{th}~\text{waiting},~i=1,...,N \\
\end{aligned}
$$

Now, 
$$
T|N = \sum_{i=1}^{N-1}(U_i + W_i) + W_N
$$
Note that for $i = 1,...,N-1$, $U_i$ and $W_i$ are iid measurements for the `short` observations with means of $\mu_{U,short}$ and $\mu_{W,short}$ respectively. $W_N$ is the length of waiting for the last `long` observation with population mean of $\mu_{W,long}$.

By Law of Total Expectation,
$$
\begin{aligned}
\mu_T
&= \mathrm{E}(T) \\
&= \mathrm{E}(\mathrm{E}(T|N)) \\
&= \mathrm{E}(\mathrm{E}(\sum_{i=1}^{N-1}(U_i + W_i) + W_N)) \\
&= \mathrm{E}(\sum_{i=1}^{N-1}(\mathrm{E}(U_i)) + \mathrm{E}(W_i)) + \mu_{W,long} 
~~~~ \text{, due to the indep.}\\
&= (\mu_{U,short} + \mu_{W,short})\mathrm{E}(N-1) + \mu_{W,long} 
~~~~ \text{, identically distr.}\\
&= (\mu_{U,short} + \mu_{W,short})(\frac{1}{p}-1) + \mu_{W,long}
\end{aligned}
$$

The parameters of interets are $p,~\mu_{U,short},~\mu_{W,short},~\mu_{W,long},\text{ and } \mu_T$.

## Computation
```{r}
p_hat <- sum(faithful$grp == "long")/dim(faithful)[1]

mu_ushort_hat <- faithful %>%
  filter(grp == "short") %>%
  select(eruptions) %>%
  unlist() %>%
  mean()

mu_wshort_hat <- faithful %>%
  filter(grp == "short") %>%
  select(waiting) %>%
  unlist() %>%
  mean()

mu_wlong_hat <- faithful %>%
  filter(grp == "long") %>%
  select(waiting) %>%
  unlist() %>%
  mean()
```

## Estimation
```{r}
mu_t_hat <- (mu_ushort_hat + mu_wshort_hat)*(1/p_hat - 1) + mu_wlong_hat
print(mu_t_hat)
```
Based on the model, if the visitor wants to see an eruption lasting more than 3 minutes, it’s estimated that he/she has to wait for `r round(mu_t_hat, 3)` under the condition of just missing the previous eruption.

